### 2019-07-17 data_analysis/data_analysis_comparison.Rmd
large C grid analysis complete

---
title: "Data Analysis Comparison"
author: "Alex Paynter"
date: "7/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(magrittr)
source("./figures_tables.R")
```

## Introduction

In this document we show the data analysis results using two choices for $C$ grid.  The small grid was our pre-specified plan, and we noticed that the $\widehat{C}$ estimates were near the upper bound of the grid for many of the methods.  This motivated the use of a larger grid to see if the results were the same.

#

## Small $C$ grid.

Let $c$ be the maximum observed richness of a sample (maximum over the replciates).  In the small $C$ grid simulation we iterate our optimization over $C = \{c, 1.2*c, 1.4*c, \dots, 20*c\}$, or in other words we set a lower bound for the coverage of our sample at 5\%.  This is the same grid we used in all simulations.  After running our algorithm with all methods we find

```{r small_out}
small_out <- process_sample_list(lakes_results,
                                 sample_names = 2009:2011,
                                 latex_format = F)
small_out
```
Here is a very preliminary plot, with the points slightly horizontally jittered:
```{r small_plot}
library(ggplot2)
# Room for improvement here:
plot_fn <- function(output_mat) {
  df <- output_mat %>%
    as.data.frame %>%
    tibble::rownames_to_column(.) %>%
    tidyr::gather(., key = "year", value = "estimate", -rowname)

  # cleaning up names:
  df["method"] <-df[["rowname"]] %>%
    stringr::str_replace(., "_", " ") %>%
    stringr::str_replace(., "m", "M") %>%
    factor(., levels = unique(.))

  pd <- position_dodge(0.1)
  ggplot(df, aes(x = year, y = estimate, group = method, color = method)) +
    geom_point(position = pd) + 
    geom_line(position = pd) +
    labs(y = "C estimate", x = "year")
}
plot_fn(small_out)
```

Given that the maximum observed richness in 2009, 2010, 2011 is `r get_cc(lakes_results)`,  this caps our $C$ estimates at `r formatC(get_cc(lakes_results) * 20, format = "d")` respecitvely.  Therefore we have many estimates which are near the arbitrarily chosen upper bound of our $C$ grid.

## Large $C$ grid

In the large $C$ grid we iterate over $C = \{c, 1.5*c, 2*c, \dots, 50*c\}$, and the results are

```{r big_out}
big_out <- process_sample_list(lakes_results_bigger_C_grid, 
                    sample_names = 2009:2011,
                    latex_format = F)
big_out

```

Here is a very preliminary plot, with the points slightly horizontally jittered:
```{r big_plot}
library(ggplot2)
# Room for improvement here:
plot_fn <- function(output_mat) {
  df <- output_mat %>%
    as.data.frame %>%
    tibble::rownames_to_column(.) %>%
    tidyr::gather(., key = "year", value = "estimate", -rowname)

  # cleaning up names:
  df["method"] <-df[["rowname"]] %>%
    stringr::str_replace(., "_", " ") %>%
    stringr::str_replace(., "m", "M") %>%
    factor(., levels = unique(.))

  pd <- position_dodge(0.1)
  ggplot(df, aes(x = year, y = estimate, group = method, color = method)) +
    geom_point(position = pd) + 
    geom_line(position = pd) +
    labs(y = "C estimate", x = "year")
}
plot_fn(big_out)
```

Our results for 2009, 2010, 2011 under this analysis are capped at `r formatC(get_cc(lakes_results) * 50, format = "d")`.  Unfortunately many of the results are still very close to the upper bound.

## Conclusions

We don't learn anything new by expanding the $C$ grid.  We might learn more by tinkering with the $\lambda^{\text{grid}}$, but I'm inclined to just report the results of the small grid which is what we did in our simulations.  We can discuss the difficulty in selecting of $\lambda^{\text{grid}}$ as a limitation.