### 2019-08-13 data_analysis/data_analysis_expanded.Rmd
Expanded data analysis update

---
title: "Data Analysis Expanded"
author: "Alex Paynter"
date: "August 9, 2019"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, include = T, cache = T, message = F, results = 'hide')
knitr::opts_knit$set(root.dir = here::here("data_analysis"))
```

```{r load_data}
library(magrittr)
library(rre)
library(dplyr);
#setwd("/Users/alex/Dropbox/classGrad/IS_Wil/rre_sims/data_analysis/")

lakes <- readr::read_tsv("./lakes_data.txt") %>% as.data.frame()
lakes_metadata <- readr::read_tsv("./lakes_metadata.txt") %>% as.data.frame()

# goal is to grab a set of samples from the same year and site
get_samples <- function(yr, zone, per = NULL, month = NULL) {
  quo_period <- dplyr::enquo(per)
  sample_list <- lakes_metadata %>%
    dplyr::filter(Site == zone) %>%
    dplyr::filter(Years == yr)
  if (!is.null(per)) {
    sample_list %<>%  dplyr::filter(Period == per) 
  }
  if (!is.null(month)) {
    sample_list %<>%  dplyr::filter(Months == month) 
  }
  sample_list %<>% dplyr::select(SampleID) %>% unlist
  tables <- lakes[,colnames(lakes) %in% sample_list] %>%
    lapply(., rre::make_frequency_count_table)
  return(tables)
}

#test <- rre::unregularized_mle(list_of_samples[[1]], 
#                               multiplier = 2,
#                               c_seq_len = 2)

report_results_2 <- function(method_output) {
  best <- method_output[["best"]]
  c_hat <- best$ccc
  m <- c_hat/(best$cc_max)
  cat(paste0("C_hat: ", c_hat, "\n\n", round(m,2), " times greater than observed richness \n\nSelected lambda: ", best$selected_lambda, "\n\n"))
}

# Note that one of the metadata files matches no lakes sample ID:  12.07.11.1
list_of_samples <- mapply(get_samples, seq(2009, 2011), "Littoral", "Summer")

big_starts = data.frame(alpha = c(0.1, 0.01, 0.01, 0.001), 
                        delta = c(0.1, 0.01, 0.0001, 0.001))
```

## Method 0

```{r s1_m0_a}
s1_m0_a <- rre::unregularized_mle(list_of_samples[[1]], 
                                  multiplier = 200,
                                  c_seq_len = 200,
                                  starts = big_starts)
```
```{r s1_m0_a_print, results = 'asis'}
s1_m0_a %>% report_results_2(.)
```
<br>
<br>

```{r s2_m0_a}
s2_m0_a <- rre::unregularized_mle(list_of_samples[[2]], 
                                  multiplier = 200,
                                  c_seq_len = 200,
                                  starts = big_starts)
```
```{r s2_m0_a_print, results = 'asis'}
s2_m0_a %>% report_results_2(.)
```
<br>
<br>

```{r s3_m0_a}
s3_m0_a <- rre::unregularized_mle(list_of_samples[[3]],
                                  multiplier = 200,
                                  c_seq_len = 200,
                                  starts = big_starts)
```
```{r s3_m0_a_print, results = 'asis'}
s3_m0_a %>% report_results_2(.)
```
<br>
<br>

<br><br>

## Method 1

### Sample 1

```{r s1_m1_a}
# big grids and few partitions to start:
set.seed(12039)
s1_m1_a <- rre::minimum_subset_distance(list_of_samples[[1]],
                                        partitions = 5,
                                        lambda_vec = seq(0,1000,by=200),
                                        multiplier = 100,
                                        c_seq_len = 20)
```
```{r s1_m1_a_print, results = 'asis'}
s1_m1_a %>% report_results_2(.)
```
<br><br>

<br><br>

### Sample 2

```{r s2_m1_a}
# big grids and few partitions to start:
set.seed(2345629)
s2_m1_a <- rre::minimum_subset_distance(list_of_samples[[2]],
                                        partitions = 5,
                                        lambda_vec = seq(0,1000,by=200),
                                        multiplier = 100,
                                        c_seq_len = 50)
```
```{r s2_m1_a_print, results = 'asis'}
s2_m1_a %>% report_results_2(.)
```
<br><br>

<br><br>

### Sample 3

```{r s3_m1_a}
# big grids and few partitions to start:
set.seed(45698)
s3_m1_a <- rre::minimum_subset_distance(list_of_samples[[3]],
                                        partitions = 5,
                                        lambda_vec = seq(0,1000,by=200),
                                        multiplier = 100,
                                        c_seq_len = 50)
```
```{r s3_m1_a_print, results = 'asis'}
s3_m1_a %>% report_results_2(.)
```
<br><br>

<br><br>

## Method 3

### Sample 1

```{r s1_m3_a}
s1_m3_a <- rre::gof_criterion(list_of_samples[[1]], 
                              lambda_vec = seq(0,1000,by=100),
                              multiplier = 200,
                              c_seq_len = 10)
```
```{r s1_m3_a_print, results = 'asis'}
s1_m3_a %>% report_results_2(.)
```
<br>
<br>

```{r s1_m3_b}
# Zero in on lambda, use a more dense C grid:
s1_m3_b <- rre::gof_criterion(list_of_samples[[1]], 
                              lambda_vec = seq(300,500,by=10),
                              multiplier = 200,
                              c_seq_len = 20)
```
```{r s1_m3_b_print, results = 'asis'}
s1_m3_b %>% report_results_2(.)
```
<br>
<br>

```{r s1_m3_c}
s1_m3_c <- rre::gof_criterion(list_of_samples[[1]], 
                              lambda_vec = seq(450,530,by=5),
                              multiplier = 200,
                              c_seq_len = 50,
                              starts = big_starts)
```
```{r s1_m3_c_print, results = 'asis'}
s1_m3_c %>% report_results_2(.)
```
<br>
<br>

```{r s1_m3_d}
# Feed in more alpha and delta, narrow lambdagrid more.
#s1_m3_d <- rre::gof_criterion(list_of_samples[[1]], 
#                              lambda_vec = seq(505,525,by=2),
#                              multiplier = 100,
#                              c_seq_len = 50,
#                              starts = data.frame(alpha = c(0.1, 0.01, 0.01, 0.001), delta = c(0.1, 0.01, 0.0001, 0.001)))
```
```{r s1_m3_d_print, results = 'asis'}
#s1_m3_d %>% report_results_2(.)
```
<br>
<br>

<br><br>

### Sample 2
```{r s2_m3_a}
s2_m3_a <- rre::gof_criterion(list_of_samples[[2]], 
                              lambda_vec = seq(0,1000,by=100),
                              multiplier = 200,
                              c_seq_len = 10)
```
```{r s2_m3_a_print, results = 'asis'}
s2_m3_a %>% report_results_2(.)
```
<br>
<br>

```{r s2_m3_b}
# Zero in on lambda, use a more dense C grid:
s2_m3_b <- rre::gof_criterion(list_of_samples[[2]], 
                              lambda_vec = seq(0,200,by=10),
                              multiplier = 200,
                              c_seq_len = 20)
```
```{r s2_m3_b_print, results = 'asis'}
s2_m3_b %>% report_results_2(.)
```
<br>
<br>

```{r s2_m3_c}
# Zero in on lambda, use a more dense C grid:
s2_m3_c <- rre::gof_criterion(list_of_samples[[2]], 
                              lambda_vec = seq(120,200,by=5),
                              multiplier = 200,
                              c_seq_len = 50)
```
```{r s2_m3_c_print, results = 'asis'}
s2_m3_c %>% report_results_2(.)
```
<br>
<br>

```{r s2_m3_d}
# Selected lambda was on the edge of previous grid, lets try again:
s2_m3_d <- rre::gof_criterion(list_of_samples[[2]], 
                              lambda_vec = seq(150,300,by=10),
                              multiplier = 100,
                              c_seq_len = 50,
                              starts = big_starts)
```
```{r s2_m3_d_print, results = 'asis'}
s2_m3_d %>% report_results_2(.)
```
<br>
<br>

<br><br>

### Sample 3
```{r s3_m3_a}
s3_m3_a <- rre::gof_criterion(list_of_samples[[3]], 
                              lambda_vec = seq(0,1000,by=100),
                              multiplier = 200,
                              c_seq_len = 10)
```
```{r s3_m3_a_print, results = 'asis'}
s3_m3_a %>% report_results_2(.)
```
<br>
<br>

```{r s3_m3_b}
# Zero in on lambda, use a more dense C grid:
s3_m3_b <- rre::gof_criterion(list_of_samples[[3]], 
                              lambda_vec = seq(0,100,by=10),
                              multiplier = 200,
                              c_seq_len = 20)
```
```{r s3_m3_b_print, results = 'asis'}
s3_m3_b %>% report_results_2(.)
```
<br>
<br>

```{r s3_m3_c}
s3_m3_c <- rre::gof_criterion(list_of_samples[[3]], 
                              lambda_vec = seq(50,130,by=10),
                              multiplier = 200,
                              c_seq_len = 50)
```
```{r s3_m3_c_print, results = 'asis'}
s3_m3_c %>% report_results_2(.)
```
<br>
<br>

```{r s3_m3_d}
# Selected lambda was on the edge of previous grid, lets try again:
s3_m3_d <- rre::gof_criterion(list_of_samples[[3]], 
                              lambda_vec = seq(100, 200, by=10),
                              multiplier = 100,
                              c_seq_len = 50,
                              starts = big_starts)
```
```{r s3_m3_d_print, results = 'asis'}
s3_m3_d %>% report_results_2(.)
```

<br><br>